Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-30 snTomas.tex
finished (another) overhaul of the discussion section   * now includes a discussion of Cosmic Weak Lensing   * highlights the quality vs quantity strong lensing argument   * reorganizes the discussion to evaluate SN problems first, then lens models

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \changeR{15 lens models developed by 8 independent groups},
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  \changeR{However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value}
and \change{6 of these models are $\geq1.5\sigma$} above the measured
magnification.  We evaluate possible causes for this mild
bias, \changeR{and reject explanations that invoke modeling errors or
misinterpretation of the SN data. We conclude that -- at least for
this single line of sight -- the most accurate models are those with
(1) stringent quality cuts for strong-lensing constraints and (2)
relatively flexible model design}. \change{A more robust evaluation of
systematic biases could be achieved with further cooperative lens
model comparisons using a larger sample of lensed Type Ia SNe as well
as cluster lensing simulations.}
\end{abstract}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using the
MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 15 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  Points with magenta outline indicate ``unblind''
models, which were generated after the magnification of the SN was
known. The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.5$.
\label{fig:lensingtest} }
\end{center}
\end{figure}

\changeR{
Table~\ref{tab:PredictedMagnifications} gives each model's predicted}
magnification and uncertainty for a source at \change{$z=1.3457$} and
at the position of \tomas.  These models represent a broad sampling of
the techniques and assumptions that can be applied to the modeling of
mass distributions in galaxy clusters. A rigorous comparison of these
disparate lens modeling techniques is beyond the scope of this
work. Table~\ref{tab:PredictedMagnifications} includes a brief
description of each model, but for a complete discussion of the
methodology the reader is referred to the listed references.

In Figure~\ref{fig:lensingtest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  This comparison shows that these 14
models are largely consistent with each other. Naively treating each
model as an independent prediction for the magnification (and ignoring
their quoted uncertainties), one would get a mean for the full sample
of \change{$\mu=2.5\pm0.4$.  This is separated from the observed SN
magnification by $\delta\mu/\mu=23\%$, which is approximately a
1$\sigma$ difference.}

The models shown in Figure~\ref{fig:lensingtest} are separated along
the $y$ direction into two broad classes, with eight so-called {\it
parametric}\footnote{Although this nomenclature is becoming standard
in the literature, it is somewhat misleading, as the pixels or grid
cells in free-form models are effectively parameters as well. Perhaps
``simply-parameterized'' would be more accurate, though we adopt the
more common usage here.}  models on top and seven {\it free-form} models
on the bottom.  Broadly speaking, the parametric models use
parameterized density distributions to describe the arrangement of
mass within the cluster, typically associating massive dark matter
halos with the positions of bright cluster member galaxies. Therefore,
parametric models rely (to varying degrees) on the assumption that the
cluster's dark matter can be described by analytic forms such as NFW
halos \citep{Navarro:1997}, or pseudo isothermal elliptical mass
distributions \citep{Kassiola:1993}.  Free-form models divide the
cluster into a grid, generally using a
multi-scale grid to get better sampling in denser regions. Each grid cell
is assigned a mass or a potential, and then the mass values are
iteratively refined to match the observed lensing
constraints. In some cases an adaptive grid is used so
that the grid spacing itself can also be modified as the model is
iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:lensingtest}.  For example,
the Williams model (using GRALE), which has the largest uncertainties
is able to account for mass sheet degeneracy by introducing a free
parameter in the reconstruction that represents an arbitrarily scaled
mass sheet, with normalization to be chosen by the genetic algorithm.
More complex degeneracies are also accounted for in this model through
the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}

Figure~\ref{fig:lensingtest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only on
strong-lensing constraints (plotted as circles), while the other four
also use weak-lensing measurements (diamonds).  Nine of the models
were constructed using only pre-HFF data (open symbols), while six
others added new multiply-imaged systems and redshifts (filled symbols).
For twelve of these models the SN magnification qualifies as a true
``blind test'', as they were completed before the SN \tomas\
magnification measurement was known.  The remaining three models
(Bradac-v2, Lam-v2, and Diego) are technically not blind, although
none of these modelers used the SN magnification as a constraint, and
the modelers did not have access to the final SN magnification value
when constructing their model.

@@

However, beyond this first-order agreement, there is a small
systematic bias apparent. All but two of the lens models
return median magnifications that are {\it higher} than the observed
value, and \change{six of the models are discrepant by more than $1.5\sigma$.}
It is important to emphasize that SN \tomas\ only samples a single
line of sight through the cluster, and this bias to higher
magnifications is minor.  Nevertheless, a systematic
shift of this nature is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but \change{nearly} universal
bias.

\subsection{Possible Errors in Supernova Interpretation}
\label{sec:SupernovaErrors}

\changeR{Before examining the lens models or the cluster, let us first
consider whether we might have misinterpreted the data on the SN
itself.}

@@

@@

\subsection{Possible Error in Lens Model Construction}
\label{sec:LensModelErrors}

\changeR{Having now ruled out misinterpetations of the SN
data, we are left to seek an explanation for the small systematic bias
by scrutinizing the lens models.  TODO: note that 
the most successful models are those that use a
stringent criteria for strong lensing constraints, which results in a
larger fraction that have spectroscopic redshifts. 
}

\subsection{Nearby Cluster Member Galaxy}
\label{sec:NearbyClusterMemberGalaxy}

The line of sight to SN \tomas\ is 5\farcs8 from a bright cluster
member galaxy, due North of the SN position (see
Figure~\ref{fig:DiscoveryImage}).  If the mass-to-light ratio (M/L)
for this galaxy were significantly different from the M/L for other
cluster member galaxies, then its proximity to the SN sight-line could
introduce a bias in the magnification.  This bias might be
particularly acute for models that hew more closely to a
simple ``light traces mass'' assumption, such as the Sharon, CATS,
GLAFIC and Zitrin-LTM models.

We tested this hypothesis using the CATS(v2) model by allowing the mass
of the nearby cluster member galaxy to vary as a free parameter in the
model. We found that the change in the SN \tomas\ magnification
prediction was less than $\Delta\mu=0.1$.  This additional dispersion
is already included in the uncertainties quoted for that model in
Table~\ref{tab:PredictedMagnifications} and
Figure~\ref{fig:lensingtest}.  Furthermore, an erroneous M/L
value for the nearby cluster member galaxy could not explain the
systematic shift of all lens models, because some do not incorporate
cluster member galaxies into their constraints at all (the free-form
Bradac, Williams and Merten models).

\subsection{Sparse Strong Lensing Constraints}
\label{sec:SparseStrongLensingConstraints}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/system3.png}
\caption{ 
The disputed component of the multiply-imaged galaxy System 3. The
original position, as given in \citet{Merten:2011} is marked with a
dashed magenta circle labeled 3.3. The solid magenta circle
labeled \33p\ shows the revised location as proposed
in \citet{Jauzac:2014c}.  Both positions are $\sim$10\arcsec\ from the
SN location.  The inset panel shows the other two components of this
multiply-imaged system, 3.1 and 3.2 (on the same scale as the main
panel), which are located north-east of the cluster center, and have a
secure spectroscopic redshift of $z=3.98\pm0.02$ \citep{Johnson:2014}.
\label{fig:system3} }
\end{center}
\end{figure}

At the outset of the HFF survey, only 33 multiply-imaged galaxies
behind Abell 2744 were known.  One might expect that the addition of
new lensing constraints from the HFF data would improve the model
constraints and reduce the tension between the predicted and observed
magnification.  However, for two of the most recent models (CATSv2,
Lam; \citealt{Jauzac:2014c,Lam:2014}) that include new strong-lensing
constraints, the predicted magnification is still significantly higher
than the observed value.  Again the CATS model family provides a
useful case study.  The CATS(v2) model \citep[from][]{Jauzac:2014c}
shares the same lens modeling software (LensTool), and the same
primary methodology as the earlier CATS models, though with two key
differences: (1) the CATS(v2) model changes the location of a
multiply-imaged system near the SN \tomas\ position; and (2) the model
incorporates 33 new multiply imaged galaxies.  These improvements
should in principle make the CATS(v2) model magnification estimates
more accurate, though we in fact see the opposite result: the
CATS(v2) model predicts a $\mu$ that is higher than the measured
value by $4.3\sigma$, and higher than either of the previous two
versions of the CATS models.

The change in location proposed by \citet{Jauzac:2014c} is for image
3.3, the third component of a triply-imaged
galaxy \citep{Merten:2011}, shown in Figure~\ref{fig:system3}.  Images
3.1 and 3.2 of this system have a secure spectroscopic redshift of
$z=3.98\pm0.02$ \citep{Johnson:2014}. \citet{Jauzac:2014c} found that
reassigning this multiple image to a source roughly 8\arcsec\ to the
south results in a tighter model reconstruction for the locations of
other multiply-imaged systems in the vicinity \citep[though][argue
against the reassignment based on the colors of the two sources in
question]{Lam:2014}.  Given the large discrepancy between the
CATS(v2) model and the SN measurement, one might question whether
this change in the position of image 3.3 is driving the increase in
the predicted $\mu$ for SN \tomas. However, \citet{Jauzac:2014c} also
examined a version of the CATS(v2) model where this single source
is left in the original position, and find that it does not
significantly affect the predicted magnification.

Furthermore, the shift to higher magnifications is not unique to the
SN \tomas\ position.  \citet{Jauzac:2014c} noted that their updated
mass model of Abell 2744 results in a systematic increase in the
magnification values across the cluster field. Examining a sample of
$\sim$30 multiple images, \citeauthor{Jauzac:2014c} found that the
magnifications increased by a factor of $\sim1.5$ relative to pre-HFF
models.  Although the single sight-line to SN \tomas\ can
only provide a very limited test of the model,
the $4.1\sigma$ discrepancy is at least a strong suggestion
that even with a large number of high-quality strong-lensing
constraints it is still possible to generate magnification
maps that are susceptible to systematic biases.

\subsection{Mass Profile Extrapolation}
\label{sec:MassProfileExtrapolation}

Eight of the mass models evaluated here are constrained only using
strong-lensing features such as multiply-imaged background galaxies
and highly magnified arcs.  There are now $\sim$150 known lensed
images behind Abell 2744 \citep{Jauzac:2014c}, but SN \tomas\ is
located several arcseconds outside the core region of the cluster
where these multiply-imaged galaxies are found.  Therefore these eight
``strong-lensing only'' models must necessarily rely on extrapolations
to provide a prediction for the SN \tomas\ magnification.  Let us
suppose that the true cluster density profile for Abell 2744 happens
to have a sharp drop right at the edge of this core strong-lensing
region. In that case these eight models might well overestimate the
mass interior to the \tomas\ position, and thus systematically
overestimate the magnification.

The other three models in our comparison set also use the strong
lensing constraints from the core region, but additionally utilize
weak lensing measurements to provide additional constraints on the
mass distribution farther from the cluster core.  The signal from weak
lensing relies on a large sample of background galaxies, so this
constraint operates principally at separations more than 1\arcmin\
from the cluster core.  At a projected separation of $\sim40\arcsec$,
SN \tomas\ falls in between the strong- and weak-lensing regimes, and
one would expect that models incorporating both of those constraints
would be less susceptible to the possible systematic bias of
a sharply steepening mass profile.  The lowest magnification estimate
in our sample (from the Merten model) is among this sub-sample.
However, we find that collectively these three strong+weak models
(shown as diamonds in Figure~\ref{fig:lensingtest}) exhibit the same
propensity to overestimate the magnification along this line of sight.
Thus, a sudden change in the mass profile outside the strong lensing
region is not a likely explanation for this small systematic bias.

With two versions of the CATS model, we also have a more direct test
of the effect of introducing weak lensing constraints.  The initial
CATS(v1) model uses only strong-lensing constraints, and gives
$\mu=2.29^{+0.05}_{-0.04}$, slightly higher than the measured value.
A second iteration of this model, labeled here as CATS(v1sw), used the
same strong-lensing features, but added in weak lensing constraints.
The revised magnification of $\mu=2.62\pm0.18$ is further from the
measured value, which serves to reinforce the conclusion that adding
weak lensing constraints cannot, by itself, resolve this bias.

\subsection{Cosmological Uncertainty}

All of the models presented here use a fixed assumption for the
parameters of a standard \LCDM\ cosmology (typically: \Om=0.3,\OL=0.7,\Ho=70 km
s$^{-1}$ Mpc$^{-1}$).  The parameter values selected for this fiducial
cosmology do have an impact on the inferred magnifications, and
variation of these parameters must be accounted for when using cluster
mass models for cosmological tests \citep{Zitrin:2014b}. 
\citep{Bayliss:2015} demonstrated that
changing the cosmological parameter within the range of current
acceptable cosmologies \citep[e.g.][]{Hinshaw:2013,Planck-Collaboration:2015},
introduces a magnification uncertainty that is comparable in magnitude
to that of the statistical uncertainties.  This leads to the
conclusion that the lens modeling uncertainties presented here are
underestimated, and should be significantly increased.

\change{
We have used the Sharon(v2) model to provide a quantitative example of
how the magnification errors are affected by inclusion of this
cosmological parameter uncertainty.  Following \citealt{Bayliss:2015},
we propagate cosmological parameter variations into the magnification
uncertainties, and find that the 68\%\ confidence range for the
magnification at $z=1.3457$ increases from [2.63,2.83] to [2.42,2.85].
The best-fit magnification value of $mu=2.74$ and the median value of
$mu=2.59$ remain the same. For this model only,
Table~\ref{tab:PredictedMagnifications} includes this cosmological
uncertainty. In this case, the Sharon(v2) model moves from being discrepant at
the $2.5\sigma$ level to a $1.7\sigma$ tension.  Incorporating this
cosmological parameter uncertainty in the other models would similarly
reduce the tension between models and observation, but would not
resolve the overall systematic shift.  Even a factor of 2.5 increase
in the uncertainties (the most drastic change seen in the analysis
of \citep{Bayliss:2015}) would still leave the most extreme outlier
(CATS(v2)) discrepant by $2.9\sigma$.  }

@@

\change{
Taking advantage of the availability of 14 well-constrained lens
models for the Abell 2744 cluster, we have used SN \tomas\ to ask how
accurately these lens models can predict the magnification along this
line of sight.  We find that these models are consistent, and fairly
accurate, collectively predicting $\mu=2.53\pm0.38$.  This is
encouraging, and reinforces the quality and value of these public lens
models for studying magnified background objects.  However, we note
that all but one of the tested models provides a median magnification
value that is larger than we observe.  This may be an indication that
there is a small systematic bias inherent to this cluster or
sight-line.
}

We have speculated on what could be the origin for such a systematic bias,
evaluating seven possible causes:

\begin{enumerate}
\item The close proximity of a cluster member galaxy is leading to a
biased magnification, because that galaxy's M/L is atypical.
\item The models are limited by a scarcity of strong lensing constraints
\item Model uncertainties are underestimated, by not accounting for 
       uncertainties in cosmological parameters.
\item The cluster mass profile exhibits a sudden change outside of the core
strong-lensing region, where we lack multiply-imaged background
sources to constrain it.
\item The redshift assigned to SN \tomas\ is in error.
\item SN \tomas\ is dimmed by foreground dust.
\item SN \tomas\ is not a normal Type Ia SN.
\end{enumerate}

\noindent The first four of these take on the discrepancy from 
the lens modeling side, but none are completely satisfactory. For each
of these scenarios, we would expect that certain subsets of the
available lens models would be less sensitive to the proposed
systematic bias.  However, the small discrepancy between the predicted
and observed lensing magnification is persistent when we bifurcate the
sample of lens models according to methodology (parametric
vs. free-form), number of strong-lensing constraints (pre- vs
post-HFF), and scope of lensing constraints (strong-lensing only
vs. strong+weak). Thus, comparison of the available models does not
provide any definitive explanation for this magnification tension.

At least for this particular sight-line, the free-form models perform
better in the sense that the mean and standard error for this subset
($2.37\pm0.35$) is closer to the measured value ($2.03\pm0.29$) than
is the parametric set ($2.59 +- 0.37$).  These free-form models also
deliver more realistic uncertainties (with the exception of the
Bradac(v2) model -- see \citet{Wang:2015} for further discussion of
the changes made from v1 to v2 of that model). \change{The free-form
model family also includes the only model that predicts a
magnification lower than the observed value (Diego).}  This may
indicate that the free-form approach in general is more successful in
capturing the uncertainty from biases inherent to the mass modeling
exercise, such as the mass sheet degeneracy and more complex
``approximate invariances'' \citep{Liesenborgs:2012,Schneider:2014}.
It may also be the case that the greater flexibility of the free-form
models makes them better suited for predicting magnifications of
sources that fall outside the strong-lensing region -- such as
SN \tomas.

The last three possible explanations presume an error in the
interpretation of the available SN data.  We reject the possibility
that a redshift error is the primary cause of the discrepancy, as the
redshift evidence is well supported by multiple lines of evidence from
both the SN and the presumed host galaxy. Changing the redshift within
the constraints of these complementary observations does not resolve
the tension. We also find it implausible that there is sufficient dust
in the cluster or elsewhere along the line of sight to account for the
magnification discrepancy.  A dust-induced bias is also disfavored by
the very blue color of SN \tomas.

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \changeR{15 lens models developed by 8 independent groups},
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  \changeR{However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value}
and \change{6 of these models are $\geq1.5\sigma$} above the measured
magnification.  We evaluate possible causes for this mild
bias, \changeR{and reject explanations that invoke modeling errors or
misinterpretation of the SN data. We conclude that -- at least for
this cluster and this line of sight -- the most accurate models are
those that employ stringent quality cuts for strong-lensing
constraints.} \change{A more robust evaluation of systematic biases
could be achieved with further cooperative lens model comparisons
using a larger sample of lensed Type Ia SNe as well as cluster lensing
simulations.}
\end{abstract}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using the
MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 15 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  Points with magenta outline indicate ``unblind''
models, which were generated after the magnification of the SN was
known. The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.5$.
\label{fig:LensingTest} }
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} gives each model's predicted
magnification and uncertainty for a source at \change{$z=1.3457$} and
at the position of \tomas.  These models represent a broad sampling of
the techniques and assumptions that can be applied to the modeling of
mass distributions in galaxy clusters.  The first eight models in this
table are so-called {\it parametric}\footnote{Although this
nomenclature is becoming standard in the literature, it is somewhat
misleading, as the pixels or grid cells in free-form models are
effectively parameters as well. Perhaps ``simply-parameterized'' would
be more accurate, though we adopt the more common usage here.}
models, and the bottom seven are {\it free-form} models.  Broadly
speaking, the parametric models use parameterized density
distributions to describe the arrangement of mass within the cluster,
typically associating massive dark matter halos with the positions of
bright cluster member galaxies. Therefore, parametric models rely (to
varying degrees) on the assumption that the cluster's dark matter can
be described by analytic forms such as NFW halos \citep{Navarro:1997},
or pseudo isothermal elliptical mass
distributions \citep[PIEMD][]{Kassiola:1993}.  Free-form models divide
the cluster into a grid, generally using a multi-scale grid to get
better sampling in denser regions. Each grid cell is assigned a mass
or a potential, and then the mass values are iteratively refined to
match the observed lensing constraints. In some cases an adaptive grid
is used so that the grid spacing itself can also be modified as the
model is iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.  
A more rigorous comparison of these
disparate lens modeling techniques is beyond the scope of this
work.  For a complete discussion of each model's
methodology, the reader is directed to the listed references.

In Figure~\ref{fig:LensingTest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  To first order, this
comparison shows that these 15 models are largely consistent with each
other and with the observed magnification of SN \tomas. Naively
treating each model as an independent prediction for the magnification
(and ignoring their quoted uncertainties), one would get a mean for
the full sample of \change{$\mu=2.5\pm0.4$.  This is separated from
the observed SN magnification by $\delta\mu/\mu=23\%$, which is
approximately a 1$\sigma$ difference.}

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:LensingTest}.  For example,
the Williams model (using GRALE) has the largest uncertainties, in
part because it accounts for mass sheet degeneracy by introducing a
free parameter in the reconstruction that represents an arbitrarily
scaled mass sheet, with normalization to be chosen by the genetic
algorithm.  More complex degeneracies are also accounted for in this
model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}

Figure~\ref{fig:LensingTest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only on
strong-lensing constraints (plotted as circles), while the other four
also use weak-lensing measurements (diamonds).  Nine of the models
were constructed using only pre-HFF data (open symbols), while six
others added new multiply-imaged systems and redshifts (filled symbols).
For twelve of these models the SN magnification qualifies as a true
``blind test'', as they were completed before the SN \tomas\
magnification measurement was known.  The remaining three models
(Bradac-v2, Lam-v2, and Diego) are technically not blind, although
none of these modelers used the SN magnification as a constraint, and
the modelers did not have access to the final SN magnification value
when constructing their model.

@@

However, beyond this first-order agreement, there is a small
systematic bias apparent. All but two of the lens models
return median magnifications that are {\it higher} than the observed
value, and \change{six of the models are discrepant by more than $1.5\sigma$.}
It is important to emphasize that SN \tomas\ only samples a single
line of sight through the cluster, and this bias to higher
magnifications is minor.  Nevertheless, a systematic
shift of this nature is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but \change{nearly} universal
bias.  \changeR{We first consider whether a misinterpretation of the data on the SN
itself can account for the observed systematic bias, and then 
 examine the lens models.}

\subsection{Possible Errors in Supernova Analysis}
\label{sec:SupernovaError}

@@

\subsection{Possible Errors in Lens Modeling}
\label{sec:LensModelComparison}

\changeR{
Having now ruled out misinterpetations of the SN data, we are left to
seek an explanation for the mild tension by scrutinizing the lens
models.  Here we evaluate four possible sources for a systematic bias to
appear in the lens models.
}

\subsubsection{Cosmic Weak Lensing}
\changeR{ 
In addition to the dominant gravitational lensing from the foreground
cluster, a distant source like SN \tomas\ will also be subject to the
cumulative weak lensing effects due to uncorrelated large scale
structure along the line of sight, sometimes called ``Cosmic Weak
Lensing'' \citep[CWL;][]{Wong:2011,Host:2012,Bayliss:2014,McCully:2014,DAloisio:2014}.
This CWL effect is most important for very high redshift sources
($z>5$), which have a much longer path length over which to encounter
large scale structure and for highly magnified sources very near to
the cluster lensing critical curve (because CWL can perturb the
position of the critical curve).  For SN \tomas\ at the modest
redshift of $z=1.35$ and far from the critical curve, CWL should
increase the scatter in the lensing magnification at a level much less
than 10\% \citep{DAloisio:2014}.  The CWL effect is therefore unable
to account for the observed discrepancy of $\sim23$\% by
itself.
}

\subsubsection{Nearby Cluster Member}
\changeR{
A component of the Abell 2744 lens that is particularly relevant to SN \tomas\ is the cluster member
galaxy that lies just 5\farcs8 north of the SN position (see
Figure~\ref{fig:DiscoveryImage}).  If the mass-to-light ratio or the
profile of the dark matter halo for this galaxy were significantly
different from other cluster member galaxies, then its
proximity to the SN sight-line might drive a bias in the magnification.
We tested this hypothesis using the CATS(v2) model by allowing the mass
of the nearby cluster member galaxy to vary as a free parameter in the
model. We found that the change in the SN \tomas\ magnification
prediction was less than $\Delta\mu=0.1$.  This additional dispersion
is already included in the uncertainties quoted for that model in
Table~\ref{tab:PredictedMagnifications} and
Figure~\ref{fig:LensingTest}.  Furthermore, an erroneous M/L
value for the nearby cluster member galaxy could not explain the
systematic shift of all lens models, because some do not incorporate
cluster member galaxies into their constraints at all (the free-form
Bradac, Williams and Merten models).
}

\subsubsection{Cosmological Parameter Uncertainty}
\changeR{
Most lens models make a
fixed assumption for the values of cosmological parameters in a
standard \LCDM\ cosmology (typically: \Om=0.3,\OL=0.7,\Ho=70 km
s$^{-1}$ Mpc$^{-1}$). This can introduce a systematic magnification
error that is comparable in magnitude to that of the statistical
uncertainties \citep{Zitrin:2014b,Bayliss:2015}.  
Incorporating this
cosmological parameter uncertainty would increase the model magnification errors typically by 1-10\%. 
This would reduce the tension between models and observation, but would not
resolve the overall systematic shift.  Even a factor of 2.5 increase
in the uncertainties (the most drastic change seen in the analysis
of \citet{Bayliss:2015}) would still leave the most extreme outlier
(CATS(v2)) discrepant by $2.9\sigma$. 
}

\subsubsection{Misidentification of a multiple image}

\changeR{
\citet{Jauzac:2014c} proposed a correction
for the location of image 3.3, the third component of a triply-imaged
galaxy \citep{Merten:2011} at $z=3.98\pm0.02$ \citep{Johnson:2014}.
This possibly specious multiple image is only $\sim$10\arcsec\ from
the SN \tomas\ position, meaning a change in its location could have a
large impact on the predicted magnification for
SN \tomas\ \citep[see][for a quantitative discussion of this
effect]{Johnson:2014}.  However, \citet{Jauzac:2014c} also evaluated a
version of the CATS(v2) model in which this single source is left in
the original position and found that the location of image 3.3 does
not significantly affect the predicted magnifications.  Furthermore,
dividing our 15 models into those using the original location vs those
that adopt the new position, we find that both groups include models
that are within 1$\sigma$ of the observed $\mu$ as well as
significantly discrepant points.
}

\subsection{Quantity and Quality of Lensing Constraints}
\label{sec:LensModelComparison}

\changeR{
None of the possibilities described above can completely account for
the mild tension between the observed and predicted magnifications.
However, in lieu of a physical explanation, we can simply take a
pragmatic approach and seek to identify any shared characteristics of
the most accurate models.  In this way SN \tomas\ may serve as a guide
for optimizing future lens models for magnification predictions.  }

@@

\changeR{
As can be seen in Table~\ref{tab:LensModels} and
Figure~\ref{fig:LensingTest}, among the 7 models that yield
magnifications discrepant by more than 1$\sigma$ we find a variety of
methodologies (both parametric and free-form) and a mix of lensing
constraints (both strong-lensing only and strong+weak).  The left
panel of Figure~\ref{fig:Nspecz} shows that this subset of 7 also
includes a wide range in the number of multiply-imaged galaxies used
as strong-lensing constraints (from $\sim$10 to 50).  Given that the
discrepancy persists across all of these categories, none of them can
provide a satisfactory explanation for the small observed bias.
}

\changeR{
However, the right panel of Figure~\ref{fig:Nspecz} shows that 3 of
the most successful lens models (Bradac(v2), Lam(v2), Diego) do share
a common quality: they all use a set of multiply-imaged galaxies for
which a large fraction have spectroscopic redshifts.  In practice,
this means that the strong-lensing constraints in these models are
defined more stringently than other models. Some other models may use
a larger set of strong-lensing features, but the dearth of
spectroscopic constraints means that it is more likely that some of
those multiply-imaged galaxies are assigned with incorrect redshifts.
}

\changeR{
This comparison suggests that to get accurate magnifications, lens
models should employ a strict vetting of the strong-lensing
constraints.  Multiply-imaged systems should ideally have
spectroscopic redshifts, or at least very well constrained photometric
redshifts.  The {\it quantity} of strong-lensing constraints is not as
important as their {\it quality}.
}

@@

\change{
Taking advantage of the availability of 15 well-constrained lens
models for the Abell 2744 cluster, we have used SN \tomas\ to ask how
accurately these lens models can predict the magnification along this
line of sight.  We find that these models are consistent, and fairly
accurate, collectively predicting $\mu=2.5\pm0.4$, within 1$\sigma$ of
the measured value.  This is encouraging, and reinforces the quality
and value of these public lens models for studying magnified
background objects.  However, we note that all but two of the tested
models provide a median magnification prediction that is larger than
the measurement from the SN, and all the models with significant
discrepancy are too high.  This may be an indication that there is a
small systematic bias at work.  }

We have speculated on what could be the origin for such a systematic bias,
first considering and rejecting three possible explanations that presume an error in the
interpretation of the available SN data.

\begin{enumerate}
\item Redshift: We reject the possibility that a redshift error is the primary cause
of the discrepancy, as the redshift evidence is well supported by
multiple lines of evidence from both the SN and the presumed host
galaxy. Changing the redshift within the constraints of these
complementary observations does not resolve the tension. 
\item Dust: We find
it implausible that there is sufficient dust in the cluster or
elsewhere along the line of sight to account for the magnification
discrepancy.  A dust-induced bias is also disfavored by the very blue
color of SN \tomas. SN \tomas\ is dimmed by foreground dust.
\item Misclassification: The combination of spectroscopic and
photometric evidence strongly supports our classification of
SN \tomas\ as a normal Type Ia SN.  The most plausible
mis-classification would be that the object is a peculiar Type Ia of
the SN 2006bt-like sub-class.  Although the light curve shape could
allow this possibility, the blue color of \tomas\ can once again
reject this alternative. 
\end{enumerate}

\noindent

\changeR{
Turning to the lens models, we have considered 4 ways in which a bias
might be introduced into the lens models:
}

\begin{enumerate}
\item Cosmic weak lensing 
\item Peculiar mass for a nearby cluster member
\item Cosmological parameter uncertainty
\item Misidentification of a multiple image
\end{enumerate}

\noindent
\changeR{
Although each of these could plausibly introduce a small systematic
bias, we find that none are likely to be significant enough
to completely resolve the tension between models and observations.
}

\changeR{
Finally, we have examined whether there is a simple prescription for
the kind of strong lensing constraints that are most likely to deliver
an accurate magnification prediction for this sight-line. 
 We find
that the number of multiply imaged systems used is not in itself
predictive of model accuracy.  Rather, it is the fraction of multiply
imaged systems that have spectroscopic redshifts that is most
correlated with model accuracy.  A reasonable interpretation 
-- at least for this cluster and this particular set of models -- is
that the spectroscopic redshift fraction serves as an effective proxy
for the ``quality'' of the strong lensing constraints. This quality of
the input data appears to be a key ingredient for deriving accurate
magnifications outside the strong-lensing regime.  }
